---
title: "Introduction_HMMCopyClusterPloidy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction_HMMCopyClusterPloidy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
# library(HMMCopyClusterPloidyv1.2)   
```

### Step1. Read the path/file of HMMcopyRead.csv.gz and process it. - Removing artifacts by High Gini coef and CNV state
```{r preprocessing}
InputDirectly <- "/Users/lees130/Library/CloudStorage/OneDrive-NYULangoneHealth/N07_DLPplus_scWGS/04_DLPplus_1qGain_8ptData/"

####### ++++++++++++  Patient 1
HMMcopyReadFile <- paste0(InputDirectly, "01_040_143929A/hmmcopy_reads.csv.gz")
MySampleID="01_040_143929A"

####### ++++++++++++  Patient 2
HMMcopyReadFile <- paste0(InputDirectly, "01_056_143929A/hmmcopy_reads.csv.gz") 
MySampleID="01_056_143929A"

####### ++++++++++++  Patient 3
HMMcopyReadFile <- paste0(InputDirectly, "01_206_143839A/hmmcopy_reads.csv.gz")
MySampleID="01_206_143839A"

####### ++++++++++++  Patient 4
HMMcopyReadFile <- paste0(InputDirectly, "01_213_143839A/hmmcopy_reads.csv.gz")
MySampleID="01_213_143839A"

####### ++++++++++++  Patient 5
HMMcopyReadFile <- paste0(InputDirectly, "01_235_143921A/hmmcopy_reads.csv.gz") 
MySampleID="01_235_143921A"

####### ++++++++++++  Patient 6
HMMcopyReadFile <- paste0(InputDirectly, "01_274_143921A/hmmcopy_reads.csv.gz") 
MySampleID="01_274_143921A"

####### ++++++++++++  Patient 7
HMMcopyReadFile <- paste0(InputDirectly, "01_282_143855A/hmmcopy_reads.csv.gz") 
MySampleID="01_282_143855A"

####### ++++++++++++  Patient 8
HMMcopyReadFile <- paste0(InputDirectly, "09_025_143855A/hmmcopy_reads.csv.gz")
MySampleID="09_025_143855A"
### Output: 587 cells

####### ++++++++++++  Patient 9
HMMcopyReadFile <- paste0(InputDirectly, "BM_CD138_143936A/hmmcopy_reads.csv.gz")
MySampleID="BM_CD138_143936A"

### Step1. Process HMMCopy CNA state data  
###       - Remove outlier bins that have NA values > 20 % of cells 
###       - Remove artifact cells with Gini coefficient > median(GiniCoefficient)+ mad(GiniCoefficient)
HMMcopy_AfterRmvCell_WithoutNABinRegion <- CNVStateProcess(HMMcopyReadFile,RemoveArtifactCell=TRUE, BinRegionDelete=TRUE, MySampleID)
dim(HMMcopy_AfterRmvCell_WithoutNABinRegion) #Pt1.  # Pt2. 881 5392  # Pt3. 656 5228  # Pt4  637 5259 # Pt5 805 5391 # Pt8  587 5413 # Pt9 589 5360


## CNVStateProcess() generates;
## - "HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_213_143839A.rds"
## - "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_235_143921A.rds" file in the Vignette folder. 
## This file has all CNA bin regions. It will be used in Step5 to make a heatmap for Chr1 centromere region heatmap. 

```

### Step2. Make a hierarchical clustering heatmap for all chromosomes.  - This heatmap still includes artifact cells of clusters.  
### This step is just to see the hierarchcial clustering with artifict cells.  But, this step is required to make ploidy plot input. 
```{r make a hierarchical clustering}

VignetteDir <- "/Users/lees130/Library/CloudStorage/OneDrive-NYULangoneHealth/N07_DLPplus_scWGS/09c_HMMCopyHierarchicalCluster_PloidyPlot_Rpackage/HMMCopyClusterPloidyv1.2/vignettes/"
### +++++++++++++ Patient 1
HMMcopy_AfterRmvCell_WithoutNABinRegion <- readRDS(paste0(VignetteDir , "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_040_143929A.rds"))
MySampleID="01_040_143929A"

### +++++++++++++ Patient 2
HMMcopy_AfterRmvCell_WithoutNABinRegion <- readRDS(paste0(VignetteDir , "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_056_143929A.rds"))
MySampleID="01_056_143929A"

### +++++++++++++ Patient 3
HMMcopy_AfterRmvCell_WithoutNABinRegion <- readRDS(paste0(VignetteDir , "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_206_143839A.rds"))
MySampleID="01_206_143839A"

### +++++++++++++ Patient 4
HMMcopy_AfterRmvCell_WithoutNABinRegion <- readRDS(paste0(VignetteDir , "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_09_025_143855A.rds"))
MySampleID="01_213_143839A"

### +++++++++++++ Patient 5
HMMcopy_AfterRmvCell_WithoutNABinRegion <- readRDS(paste0(VignetteDir , "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_235_143921A.rds"))
MySampleID="01_235_143921A"
ClusterSize=9

### +++++++++++++ Patient 6
HMMcopy_AfterRmvCell_WithoutNABinRegion <- readRDS(paste0(VignetteDir , "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_274_143921A.rds"))
MySampleID="01_274_143921A"
ClusterSize=9

### +++++++++++++ Patient 7
HMMcopy_AfterRmvCell_WithoutNABinRegion <- readRDS(paste0(VignetteDir , "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_282_143855A.rds"))
MySampleID="01_282_143855A"
ClusterSize=9

### +++++++++++++ Patient 8
HMMcopy_AfterRmvCell_WithoutNABinRegion <- readRDS(paste0(VignetteDir , "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_09_025_143855A.rds"))
MySampleID="09_025_143855A"
ClusterSize=4

### +++++++++++++ Patient 9
MySampleID="BM_CD138_143936A"
ClusterSize=4

while (!is.null(dev.list()))  dev.off()
HierarchialClusteringHeatmap(Processed_HMMcopyStateData=HMMcopy_AfterRmvCell_WithoutNABinRegion, MySampleID, AllCell="WithoutArtifact", ChromosomeToSubset="All",ClusterSize=9)

### This will generate two files 
# "CompHeatmap_CNV_WithoutArtifact_AllChr_01_206_143839A_9Cluster.pdf"
# "RowSplitHeatmapDraw_WithoutArtifact_AllChr_01_206_143839A.rds"

## error 
#. Error in dev.off() :   Run 'dev.off()' 
# annot open compressed file '/Users/lees130/Library/CloudStorage/OneDrive-NYULangoneHealth/N07_DLPplus_scWGS/09c_HMMCopyHierarchicalCluster_PloidyPlot_Rpackage/HMMCopyClusterPloidyv1.2/.Rproj.user/shared/notebooks/6586349E-Introduction_HMMCopyClusterPloidy/1/65E26DDE8743a310/cv3u6mjw7iap4_t/35eca20294054de198578c8a571a662a.snapshot', probable reason 'No such file or directory'

```


### Step3. Remove artifact cells by CNA state Low or High: Cells have CNA state 0 or 1 in 20% of total bins (n=5259) and Cells have CNA state ≥ 5 in more than 10 bins 
###         Remove clusters that have cells less than 10.
###        Then, try to make a hierarchical clustering.  KM_ClusterSize is determined after looking at the heatmap in Step2.   

```{r remove artifact cells}
VignetteDirectory <- "/Users/lees130/Library/CloudStorage/OneDrive-NYULangoneHealth/N07_DLPplus_scWGS/09c_HMMCopyHierarchicalCluster_PloidyPlot_Rpackage/HMMCopyClusterPloidyv1.2/vignettes/"

## 1st patient, 01_040_143929A
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_040_143929A.rds")
KM_ClusterSize=5
MySampleID <- "01_040_143929A"
ThresholdLowState=1; ThresholdHighState=50 # default 10.   ThresholdLowState=1 means I don't remove any cells. 
RemoveArtifactCellHeatmap(CNVStateData_AfterRmvHighGiniMeanFile, MySampleID, KM_ClusterSize, ThresholdLowState, ThresholdHighState) 

## 2nd patient, 01_056_143929A
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_056_143929A.rds")
KM_ClusterSize=6
MySampleID <- "01_056_143929A"
ThresholdLowState=1; ThresholdHighState=10 # default 10
RemoveArtifactCellHeatmap(CNVStateData_AfterRmvHighGiniMeanFile, MySampleID, KM_ClusterSize, ThresholdLowState, ThresholdHighState) 

## 3rd patients, 01_206_143839A
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_206_143839A.rds")
KM_ClusterSize=6
ThresholdLowState=1; ThresholdHighState=10 # default 10
MySampleID <- "01_206_143839A"
RemoveArtifactCellHeatmap(CNVStateData_AfterRmvHighGiniMeanFile, MySampleID, KM_ClusterSize, ThresholdLowState, ThresholdHighState) 

## 4th patients, 01_213_143839A
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_213_143839A.rds")
KM_ClusterSize <- 7;
MySampleID <- "01_213_143839A"
ThresholdLowState=0.2;
RemoveArtifactCellHeatmap(CNVStateData_AfterRmvHighGiniMeanFile, MySampleID, KM_ClusterSize, ThresholdLowState, ThresholdHighState) 

## 5th patients, 01_235_143921A
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_235_143921A.rds")
KM_ClusterSize <- 9;
MySampleID <- "01_235_143921A"
ThresholdLowState=0.18;ThresholdHighState=10 # default 10
RemoveArtifactCellHeatmap(CNVStateData_AfterRmvHighGiniMeanFile, MySampleID, KM_ClusterSize, ThresholdLowState, ThresholdHighState) 

## 6th patients, 01_274_143921A
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_274_143921A.rds")
KM_ClusterSize <- 6;
MySampleID <- "01_274_143921A"
ThresholdLowState=0.20;
RemoveArtifactCellHeatmap(CNVStateData_AfterRmvHighGiniMeanFile, MySampleID, KM_ClusterSize, ThresholdLowState, ThresholdHighState) 

## 7th patients, 01_282_143855A
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_282_143855A.rds")
KM_ClusterSize <- 7;
MySampleID <- "01_282_143855A"
ThresholdLowState=0.20;
RemoveArtifactCellHeatmap(CNVStateData_AfterRmvHighGiniMeanFile, MySampleID, KM_ClusterSize, ThresholdLowState, ThresholdHighState) 

## 8th patients, 09_025_143855A
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_09_025_143855A.rds")
KM_ClusterSize <- 5;
MySampleID <- "09_025_143855A"
ThresholdLowState=0.18;
RemoveArtifactCellHeatmap(CNVStateData_AfterRmvHighGiniMeanFile, MySampleID, KM_ClusterSize, ThresholdLowState, ThresholdHighState) 

## 9th patients, BM_CD138_143936A
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_BM_CD138_143936A.rds")
KM_ClusterSize <- 10;
MySampleID <- "BM_CD138_143936A"
ThresholdLowState=0.25;
RemoveArtifactCellHeatmap(CNVStateData_AfterRmvHighGiniMeanFile, MySampleID, KM_ClusterSize, ThresholdLowState, ThresholdHighState) 

### This generates: 
## CompHmapCNV_RmvByGiniMeanLowHighState_AllChr_09_025_143855A_LowSt0.18_527Cells_5clu.pdf
## "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_",MySampleID,"_",KM_ClusterSize,"clu.rds"
## "ClusterNumber_15LowCell_01_235_143921A_9_8_6_4.txt" This one tells which clusters of cells < 15 were removed. 

```

### Step4. Check the heatmaps made in Step3. Remove normal cluster cells.  Make a final hierarchical clustering heatmap for all chromosome,
### and chromosome1. Stack the clusters by clone order. 

```{r Remove normal cells and Reorder Cluster}
## 1st patients, 01_040_143929A
HierarchialClusterInfoFile <- paste0(VignetteDirectory, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_040_143929A_5clu.rds")
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvCellByGiniMean_01_040_143929A_LowHighState1.rds")
MySampleID <- "01_040_143929A"
ClusterNumbToRmv=NULL
ClusterReorder <- c(1,2,4,3,5)

## 2nd patients, 01_056_143929A
HierarchialClusterInfoFile <- paste0(VignetteDirectory, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_056_143929A_6clu.rds")
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvCellByGiniMean_01_056_143929A_LowHighState1.rds")
MySampleID <- "01_056_143929A"
ClusterNumbToRmv=NULL
ClusterReorder <- c(1,2,4,3,5,6)

## 3rd patients, 01_206_143839A
HierarchialClusterInfoFile <- paste0(VignetteDirectory, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_206_143839A_6clu.rds")
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvCellByGiniMean_01_206_143839A_LowHighState1.rds")
MySampleID <- "01_206_143839A"
ClusterNumbToRmv=6
ClusterReorder <- c(1,3,5,4,2)

## 4th patients, 01_213_143839A
HierarchialClusterInfoFile <- paste0(VignetteDirectory, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_213_143839A_7clu.rds")
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvCellByGiniMean_01_213_143839A_LowHighState0.2.rds")
MySampleID <- "01_213_143839A"
ClusterNumbToRmv=c(1,2) # c(1,7)
ClusterReorder <- c(5,7,6,3,4)

## 5th patients, 01_235_143921A
HierarchialClusterInfoFile <- paste0(VignetteDirectory, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_235_143921A_9clu.rds")
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvCellByGiniMean_01_235_143921A_LowHighState0.18.rds")
MySampleID <- "01_235_143921A"
ClusterNumbToRmv=c(1,4)
ClusterReorder <- c(3,2,7,5,6)

## 6th patients, 01_274_143921A
HierarchialClusterInfoFile <- paste0(VignetteDirectory, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_274_143921A_6clu.rds")
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvCellByGiniMean_01_274_143921A_LowHighState0.2.rds")
MySampleID <- "01_274_143921A"
ClusterNumbToRmv=c(1,6,5)
ClusterReorder <- c(4,3,2)

## 7th patients, 01_282_143855A
HierarchialClusterInfoFile <- paste0(VignetteDirectory, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_282_143855A_7clu.rds")
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvCellByGiniMean_01_282_143855A_LowHighState0.2.rds")
MySampleID <- "01_282_143855A"
ClusterNumbToRmv=c(1,6) # on Feb.11th, we decidied to remove cluster #6 
ClusterReorder <- c(5,4,7,2,3)

## 8th patients, 09_025_143855A
HierarchialClusterInfoFile <- paste0(VignetteDirectory, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_09_025_143855A_5clu.rds")
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvCellByGiniMean_09_025_143855A_LowHighState0.18.rds")
MySampleID <- "09_025_143855A"
ClusterNumbToRmv=NULL
ClusterReorder <- c(1,2,4,3,5)

## 9th patients, BM_CD138_143936A
HierarchialClusterInfoFile <- paste0(VignetteDirectory, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_BM_CD138_143936A_10clu.rds")
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvCellByGiniMean_BM_CD138_143936A_LowHighState0.25.rds")
MySampleID <- "BM_CD138_143936A"
ClusterNumbToRmv=c(10,8) # 6 is automatically removed.
ClusterReorder <- c(  3,5,7, 9,   4,6,1,2)

RemoveNormalClusterReorderHeatmap(HierarchialClusterInfoFile, CNVStateData_AfterRmvHighGiniMeanFile,MySampleID,ClusterNumbToRmv, ClusterReorder)

## It will generate
# HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_MySampleID.rds
# CompHmapCNV_RmvByGiniMeanLowHighState_LowSt_AllChr_01_213_143839A_389cells_RordClone.pdf
# CompHmapCNV_RmvByGiniMeanLowHighState_LowSt_Chr1only_01_213_143839A_389cells_RordClone.pdf

## ClusterNumber_15LowCell_01_056_143929A_6.txt.  Check this file and find the number of cluster to remove. 

```

### Step5. Heatmap for chromosome 1 pericentromeric region
```{r chr1 centromere region heatmap} 
## Pt. #1
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File<-paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_040_143929A.rds")
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_040_143929A.rds")
MySampleID <- "01_040_143929A"
Chr1Centromere_Heatmap(HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File, HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile, MySampleID)

## Pt. #2
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File<-paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_056_143929A.rds")
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_056_143929A.rds")
MySampleID <- "01_056_143929A"
Chr1Centromere_Heatmap(HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File, HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile, MySampleID)

## Pt. #3
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File<-paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_206_143839A.rds")
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_206_143839A.rds")
MySampleID <- "01_206_143839A"
Chr1Centromere_Heatmap(HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File, HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile, MySampleID)

## Pt. #4
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File<-paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_213_143839A.rds")
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_213_143839A.rds")
MySampleID <- "01_213_143839A"
Chr1Centromere_Heatmap(HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File, HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile, MySampleID)

## Pt. #5
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File<-paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_235_143921A.rds")
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_235_143921A.rds")
MySampleID <- "01_235_143921A"
Chr1Centromere_Heatmap(HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File, HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile, MySampleID)

## Pt. #6
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File<-paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_274_143921A.rds")
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_274_143921A.rds")
MySampleID <- "01_274_143921A"
Chr1Centromere_Heatmap(HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File, HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile, MySampleID)

## Pt. #7 patient, 01_282_143855A
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File<-paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_282_143855A.rds")
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_282_143855A.rds")
MySampleID <- "01_282_143855A"
Chr1Centromere_Heatmap(HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File, HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile, MySampleID)

## 8th patient, 09_025_143855A
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File<-paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_09_025_143855A.rds")
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_09_025_143855A.rds")
MySampleID <- "09_025_143855A"
Chr1Centromere_Heatmap(HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File, HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile, MySampleID)

## 9th patient, BM_CD138_143936A
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File<-paste0(VignetteDirectory, "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_BM_CD138_143936A.rds")
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile <- paste0(VignetteDirectory, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_BM_CD138_143936A.rds")
MySampleID <- "BM_CD138_143936A"
Chr1Centromere_Heatmap(HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File, HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile, MySampleID)



# Chr1Centromere_Heatmap(HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File, HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile, MySampleID)

### Output file name: CompHeatmap_Chr1CentromereRegion_01_235_143921A_OOOcells.pdf

```
### Step6. Ploidy plot for all chromosomes. 
```{r chr1 Ploidy plot} 
library(regioneR)
library(karyoploteR)
library(CopyNumberPlots)


DLPpHMMCopyOutDir <- "/Users/lees130/Library/CloudStorage/OneDrive-NYULangoneHealth/N07_DLPplus_scWGS/04_DLPplus_1qGain_8ptData"
VignetteDirectory <- "/Users/lees130/Library/CloudStorage/OneDrive-NYULangoneHealth/N07_DLPplus_scWGS/09c_HMMCopyHierarchicalCluster_PloidyPlot_Rpackage/HMMCopyClusterPloidyv1.2/vignettes/"

### Patient #1 - 01_040_143929A
CNVSegmentFile=paste0(DLPpHMMCopyOutDir,"/01_040_143929A/hmmcopy_segments.csv.gz");
HierarchialClusterInfoFile=paste0(VignetteDirectory, "RowSplitHeatmapDraw_WithoutArtifact_AllChr_01_040_143929A.rds");
HierarchialClusterInfoAfRmvClusterFile=paste0(VignetteDirectory, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_040_143929A_5clu.rds");
# HierarchialClusterInfoAfRmvClusterChr1File=paste0(VignetteDirectory, "RowSplitHeatmapDraw_WithoutArtifact_OnlyChr1_ByClusterSize6_01_040_143929A.rds");
CNVStateData_AfterRmvHighGiniMeanFile=paste0(VignetteDirectory, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_040_143929A.rds" );   #"/HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_040_143929A.rds");
ClusterNumbToRmv=NULL;  ClusterNumbToPick=NULL; MySampleID="01_040_143929A"

### Patient #2 - 01_056_143929A
CNVSegmentFile=paste0(DLPpHMMCopyOutDir,"/01_056_143929A/hmmcopy_segments.csv.gz");
HierarchialClusterInfoFile=paste0(VignetteDirectory, "RowSplitHeatmapDraw_WithoutArtifact_AllChr_01_056_143929A.rds");
HierarchialClusterInfoAfRmvClusterFile=paste0(VignetteDirectory, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_056_143929A_6clu.rds");
# HierarchialClusterInfoAfRmvClusterChr1File=paste0(VignetteDirectory, "RowSplitHeatmapDraw_WithoutArtifact_OnlyChr1_ByClusterSize6_01_056_143929A.rds");
CNVStateData_AfterRmvHighGiniMeanFile=paste0(VignetteDirectory, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_056_143929A.rds" );   #"/HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_056_143929A.rds");
ClusterNumbToRmv=NULL;  ClusterNumbToPick=NULL; MySampleID="01_056_143929A"

### Patient #3 - 01_206_143839A
CNVSegmentFile=paste0(DLPpHMMCopyOutDir,"/01_206_143839A/hmmcopy_segments.csv.gz");
HierarchialClusterInfoFile=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifact_AllChr_01_206_143839A.rds");
HierarchialClusterInfoAfRmvClusterFile=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_206_143839A_6clu.rds");
# HierarchialClusterInfoAfRmvClusterChr1File=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifact_OnlyChr1_ByClusterSize6_01_206_143839A.rds");
CNVStateData_AfterRmvHighGiniMeanFile=paste0(VignetteDirectory, "/HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_206_143839A.rds" );   #"/HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_206_143839A.rds");
ClusterNumbToRmv=NULL;  ClusterNumbToPick=NULL; MySampleID="01_206_143839A"

### Patient #4 - 01_213_143839A
CNVSegmentFile=paste0(DLPpHMMCopyOutDir,"/01_213_143839A/hmmcopy_segments.csv.gz");
HierarchialClusterInfoFile=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifact_AllChr_01_213_143839A.rds");
HierarchialClusterInfoAfRmvClusterFile=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_213_143839A_7clu.rds");
# HierarchialClusterInfoAfRmvClusterChr1File=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifact_OnlyChr1_ByClusterSize6_01_213_143839A.rds");
CNVStateData_AfterRmvHighGiniMeanFile=paste0(VignetteDirectory, "/HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_213_143839A.rds" );   #"/HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_213_143839A.rds");
ClusterNumbToRmv=NULL;  ClusterNumbToPick=NULL; MySampleID="01_213_143839A"


### Patient #5 - 01_235_143921A
CNVSegmentFile=paste0(DLPpHMMCopyOutDir,"/01_235_143921A/hmmcopy_segments.csv.gz");
HierarchialClusterInfoFile=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifact_AllChr_01_235_143921A.rds");
HierarchialClusterInfoAfRmvClusterFile=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_235_143921A_9clu.rds");
# HierarchialClusterInfoAfRmvClusterChr1File=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifact_OnlyChr1_ByClusterSize6_09_025_143855A.rds");
CNVStateData_AfterRmvHighGiniMeanFile=paste0(VignetteDirectory, "/HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_235_143921A.rds" );   #"/HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_235_143921A.rds");
ClusterNumbToRmv=NULL;  ClusterNumbToPick=NULL; MySampleID="01_235_143921A"

### Patient #6 - 01_274_143921A
CNVSegmentFile=paste0(DLPpHMMCopyOutDir,"/01_274_143921A/hmmcopy_segments.csv.gz");
HierarchialClusterInfoFile=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifact_AllChr_01_274_143921A.rds");
HierarchialClusterInfoAfRmvClusterFile=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_274_143921A_6clu.rds");
# HierarchialClusterInfoAfRmvClusterChr1File=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifact_OnlyChr1_ByClusterSize6_01_274_143921A.rds");
CNVStateData_AfterRmvHighGiniMeanFile=paste0(VignetteDirectory, "/HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_274_143921A.rds" );   #"/HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_274_143921A.rds");
ClusterNumbToRmv=NULL;  ClusterNumbToPick=NULL; MySampleID="01_274_143921A"

### Patient #7 - 01_282_143855A
CNVSegmentFile=paste0(DLPpHMMCopyOutDir,"/01_282_143855A/hmmcopy_segments.csv.gz");
HierarchialClusterInfoFile=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifact_AllChr_01_282_143855A.rds");
HierarchialClusterInfoAfRmvClusterFile=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_282_143855A_7clu.rds");
# HierarchialClusterInfoAfRmvClusterChr1File=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifact_OnlyChr1_ByClusterSize6_01_274_143921A.rds");
CNVStateData_AfterRmvHighGiniMeanFile=paste0(VignetteDirectory, "/HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_282_143855A.rds" );   
ClusterNumbToRmv=NULL;  ClusterNumbToPick=NULL; MySampleID="01_282_143855A"

### Patient #8 - 09_025_143855A
CNVSegmentFile=paste0(DLPpHMMCopyOutDir,"/09_025_143855A/hmmcopy_segments.csv.gz");
HierarchialClusterInfoFile=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifact_AllChr_09_025_143855A.rds");
HierarchialClusterInfoAfRmvClusterFile=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_09_025_143855A_5clu.rds");
# HierarchialClusterInfoAfRmvClusterChr1File=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifact_OnlyChr1_ByClusterSize6_01_274_143921A.rds");
CNVStateData_AfterRmvHighGiniMeanFile=paste0(VignetteDirectory, "/HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_09_025_143855A.rds" );   
ClusterNumbToRmv=NULL;  ClusterNumbToPick=NULL; MySampleID="09_025_143855A"


### Patient #9 - BM_CD138_143936A
CNVSegmentFile=paste0(DLPpHMMCopyOutDir,"/BM_CD138_143936A/hmmcopy_segments.csv.gz");
HierarchialClusterInfoFile=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifact_AllChr_BM_CD138_143936A.rds");
HierarchialClusterInfoAfRmvClusterFile=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_BM_CD138_143936A_10clu.rds");
# HierarchialClusterInfoAfRmvClusterChr1File=paste0(VignetteDirectory, "/RowSplitHeatmapDraw_WithoutArtifact_OnlyChr1_ByClusterSize6_01_274_143921A.rds");
CNVStateData_AfterRmvHighGiniMeanFile=paste0(VignetteDirectory, "/HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_BM_CD138_143936A.rds" );   
ClusterNumbToRmv=NULL;  ClusterNumbToPick=NULL; MySampleID="BM_CD138_143936A"



library(regioneR)
library(karyoploteR)
library(CopyNumberPlots)
data.panel <-1

Karyoplote_BySegment(CNVSegmentFile=CNVSegmentFile,HierarchialClusterInfoFile=HierarchialClusterInfoFile,
                     HierarchialClusterInfoAfRmvClusterFile=HierarchialClusterInfoAfRmvClusterFile,HierarchialClusterInfoAfRmvClusterChr1File=NULL,
                     CNVStateData_AfterRmvHighGiniMeanFile=CNVStateData_AfterRmvHighGiniMeanFile,
                     CNVStateData_AfterRmvHighGiniMeanClusterFile=CNVStateData_AfterRmvHighGiniMeanClusterFile,
                     ClusterNumbToRmv,  ClusterNumbToPick, ClusterNumbToPick_InChr1=ClusterNumbToPick_InChr1,MySampleID=MySampleID )






```


### Step7. Subclone phylogenetic tree
```{r phylogenetic tree}
## Pt3, 01_206  clone evolution
TreeData <- data.frame("sample" = c("01_206","01_206","01_206","01_206","01_206"), #name of sample, useful to read in all samples at once if processing multiple
                           "from" = c("x",   "Clone1", "x",      "Clone3","Clone4"), #parental Clones. 'x' is a virtual normal. 
                           "to"= c("Clone1", "Clone2", "Clone3", "Clone4","Clone5"))  #child Clone
                           #"score" = c() #a metric of CNV burden 
# Pt3 01_206 cell numbers 
NumbCellData <- data.frame("sample" = c("01_206","01_206", "01_206", "01_206","01_206","01_206"), #name of sample, useful to read in all samples at once if processing multiple
                                  "name" = c("x", "Clone1", "Clone2", "Clone3", "Clone4", "Clone5"), #Clone
                                  "counts"= c(0,19, 133,59,415, 25)) #number of cells in Clone
MySampleID <- "01_206";


## Make Phylogenetic tree
PhylogeneticTree(df=TreeData, num_cells_in=NumbCellData, MySampleID=MySampleID) 

## This will output "01_206_PhylogeneticTree.pdf" file. 

```
