---
title: "Introduction_HMMCopyClusterPloidy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction_HMMCopyClusterPloidy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
# library(HMMCopyClusterPloidyv1.2)   
```

### Step1. Read the path/file of HMMcopyRead.csv.gz and process it. - Removing artifacts by High Gini coef and CNV state
###         Remove artifact cells by CNA state Low or High: Cells have CNA state 0 or 1 in 20% of total bins (n=5259) and Cells have CNA state ≥ 5 in more than 10 bins 
###         Remove clusters that have cells less than 10.
```{r preprocessing}
InputDirectly <- "/Users/lees130/Library/CloudStorage/OneDrive-NYULangoneHealth/N07_DLPplus_scWGS/04a_DLPplus_1qGain_8ptData/"

# ####### ++++++++++++  Patient 1
# HMMcopyReadFile <- paste0(InputDirectly, "01_040_143929A/hmmcopy_reads.csv.gz")
# MySampleID="01_040_143929A"; ThresholdLowState=1; ThresholdHighState=50
# 
# ####### ++++++++++++  Patient 2
# HMMcopyReadFile <- paste0(InputDirectly, "01_056_143929A/hmmcopy_reads.csv.gz") 
# MySampleID="01_056_143929A";  ThresholdLowState=1; ThresholdHighState=50

####### ++++++++++++  Patient 3
HMMcopyReadFile <- paste0(InputDirectly, "01_206_143839A/hmmcopy_reads.csv.gz")
MySampleID="01_206_143839A";  ThresholdLowState=1; ThresholdHighState=10

####### ++++++++++++  Patient 4
HMMcopyReadFile <- paste0(InputDirectly, "01_213_143839A/hmmcopy_reads.csv.gz")
MySampleID="01_213_143839A"; ThresholdLowState=0.2; ThresholdHighState=10

####### ++++++++++++  Patient 5
HMMcopyReadFile <- paste0(InputDirectly, "01_235_143921A/hmmcopy_reads.csv.gz") 
MySampleID="01_235_143921A"; ThresholdLowState=0.18; ThresholdHighState=10

# ####### ++++++++++++  Patient 6
# HMMcopyReadFile <- paste0(InputDirectly, "01_274_143921A/hmmcopy_reads.csv.gz") 
# MySampleID="01_274_143921A"
# 
# ####### ++++++++++++  Patient 7
# HMMcopyReadFile <- paste0(InputDirectly, "01_282_143855A/hmmcopy_reads.csv.gz") 
# MySampleID="01_282_143855A"

####### ++++++++++++  Patient 8
HMMcopyReadFile <- paste0(InputDirectly, "09_025_143855A/hmmcopy_reads.csv.gz")
MySampleID="09_025_143855A"; ThresholdLowState=0.18; ThresholdHighState=10;
### Output: 587 cells

####### ++++++++++++  Patient 9
HMMcopyReadFile <- paste0(InputDirectly, "BM_CD138_143936A/hmmcopy_reads.csv.gz")
MySampleID="BM_CD138_143936A"; ThresholdLowState=0.25; ThresholdHighState=10;

### Step1. Process HMMCopy CNA state data  
###       - Remove outlier bins that have NA values > 20 % of cells 
###       - Remove artifact cells with Gini coefficient > median(GiniCoefficient)+ mad(GiniCoefficient)
HMMcopy_AfterRmvCell_WithoutNABinRegion <- CNVStateProcess(HMMcopyReadFile, RemoveArtifactCell=TRUE, BinRegionDelete=TRUE, MySampleID, ThresholdLowState, ThresholdHighState)
dim(HMMcopy_AfterRmvCell_WithoutNABinRegion) #Pt1.  # Pt2. 881 5392  # Pt3. 656 5228  # Pt4  637 5259 # Pt5 805 5391 # Pt8 09_025_143855A 519 5051 # Pt9 589 5360


## CNVStateProcess() generates;
## "HMMcopy_StateIdeal_RmvCellByGiniMean_01_040_143929A_LowHighState1.rds"  - with bin region annotation. 
## "HMMcopy_StateIdeal_RmvCellByGiniMeanLowHighState_NoBinRegionAnnot_01_040_143929A.rds"
## This file has all CNA bin regions. It will be used in Step5 to make a heatmap for Chr1 centromere region heatmap. 

## - "HMMcopy_StateIdeal_RmvCellByGiniMeanLowHighState_WithoutBinRegionAnnot_09_025_143855A.rds" file in the Vignette folder. 

```

### Step2. Run PCA and Kmean.  Make a hierarchical clustering heatmap for all chromosomes.  - This heatmap still includes artifact or low quality cells of clusters.  
###       Give 4 or 5 KmeanClusterSize and see the clusters first. You can choose which cluster to remove and how to order clusters. 
### This step is just to see the hierarchcial clustering with artifict cells.  But, this step is required to make ploidy plot input. 
```{r make a hierarchical clustering}

VignetteDir <- "/Users/lees130/Library/CloudStorage/OneDrive-NYULangoneHealth/N07_DLPplus_scWGS/09d_HMMCopy_PCAKMeanClustering_Rpackage/HMMCopyPCAKMeanPloidy/vignettes/"

# ### +++++++++++++ Patient 1         <+++++++ Heatmap is really dirty
# Processed_HMMcopyStateData <- readRDS(paste0(VignetteDir , "HMMcopy_StateIdeal_RmvCellByGiniMeanLowHighState_NoBinRegionAnnot_01_040_143929A.rds"))
# MySampleID="01_040_143929A"; KmeanClusterSize=5
# 
# ### +++++++++++++ Patient 2
# Processed_HMMcopyStateData <- readRDS(paste0(VignetteDir , "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_056_143929A.rds"))
# MySampleID="01_056_143929A"

### +++++++++++++ Patient 3
Processed_HMMcopyStateData <- readRDS(paste0(VignetteDir , "HMMcopy_StateIdeal_RmvCellByGiniMeanLowHighState_NoBinRegionAnnot_01_206_143839A.rds"))
MySampleID="01_206_143839A";  KmeanClusterSize=5

### +++++++++++++ Patient 4
Processed_HMMcopyStateData <- readRDS(paste0(VignetteDir , "HMMcopy_StateIdeal_RmvCellByGiniMeanLowHighState_NoBinRegionAnnot_01_213_143839A.rds"))
MySampleID="01_213_143839A"; KmeanClusterSize=8
# group1 group2 group3 group4 group5 group6 group7 group8 
#    123     18     17     48    173     11      7     97 

### +++++++++++++ Patient 5
Processed_HMMcopyStateData <- readRDS(paste0(VignetteDir , "HMMcopy_StateIdeal_RmvCellByGiniMeanLowHighState_NoBinRegionAnnot_01_235_143921A.rds"))
MySampleID="01_235_143921A"; KmeanClusterSize=6

# ### +++++++++++++ Patient 6
# Processed_HMMcopyStateData <- readRDS(paste0(VignetteDir , "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_274_143921A.rds"))
# MySampleID="01_274_143921A"
# ClusterSize=9
# 
# ### +++++++++++++ Patient 7
# Processed_HMMcopyStateData <- readRDS(paste0(VignetteDir , "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_282_143855A.rds"))
# MySampleID="01_282_143855A"
# ClusterSize=9

### +++++++++++++ Patient 8
Processed_HMMcopyStateData <- readRDS(paste0(VignetteDir , "HMMcopy_StateIdeal_RmvCellByGiniMeanLowHighState_NoBinRegionAnnot_09_025_143855A.rds"))
MySampleID="09_025_143855A"; KmeanClusterSize=5

### +++++++++++++ Patient 9
Processed_HMMcopyStateData <- readRDS(paste0(VignetteDir , "HMMcopy_StateIdeal_RmvCellByGiniMeanLowHighState_NoBinRegionAnnot_BM_CD138_143936A.rds"))
MySampleID="BM_CD138_143936A"; KmeanClusterSize=10



while (!is.null(dev.list()))  dev.off()

PCAKmeanClustering(Processed_HMMcopyStateData,  MySampleID=MySampleID, KmeanClusterSize)

### This will generate two files 
# "CompHmapCNV_RmvByGiniMeanLowHighState_AllChr_01_040_143929A_242Cells_5clu.pdf"  Overall heatmap with clusters. Still normal or low quality cell of cluster. 
# "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_040_143929A_5clu.rds" ## It is a list. Names are cluster number and elements are row numbers. 
# ClusterNumber_10LowCell_01_040_143929A_.txt ## show th cluster number that has < 10 cells and excluded 

## error 
#. Error in dev.off() :   Run 'dev.off()' 
# annot open compressed file '/Users/lees130/Library/CloudStorage/OneDrive-NYULangoneHealth/N07_DLPplus_scWGS/09c_HMMCopyHierarchicalCluster_PloidyPlot_Rpackage/HMMCopyClusterPloidyv1.2/.Rproj.user/shared/notebooks/6586349E-Introduction_HMMCopyClusterPloidy/1/65E26DDE8743a310/cv3u6mjw7iap4_t/35eca20294054de198578c8a571a662a.snapshot', probable reason 'No such file or directory'

## Memory issue error: 
# Error: C stack usage  7956112 is too close to the limit

```

### Step3. You got hierarchical clustering heatmap idea in step2. Remove normal or low quality cell clusters and make a final heatmap. 
###         This is the final heatmap for all chromosomes and and chromosome1. Stack the clusters by clone order. 
```{r Remove normal cells and Reorder Cluster}

# ## 1st patients, 01_040_143929A
# HierarchialClusterInfoFile <- paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_040_143929A_5clu.rds")
# CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_RmvCellByGiniMean_01_040_143929A_LowHighState1.rds")
# MySampleID <- "01_040_143929A"
# ClusterNumbToRmv=NULL
# ClusterReorder <- c(1,2,4,3,5)
# 
# ## 2nd patients, 01_056_143929A
# HierarchialClusterInfoFile <- paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_056_143929A_6clu.rds")
# CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_RmvCellByGiniMean_01_056_143929A_LowHighState1.rds")
# MySampleID <- "01_056_143929A"
# ClusterNumbToRmv=NULL
# ClusterReorder <- c(1,2,4,3,5,6)

## 3rd patients, 01_206_143839A
HierarchialClusterInfoFile <- paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_206_143839A_5clu.rds")
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_RmvCellByGiniMeanLowHighState_Kmean_01_206_143839A.rds")
MySampleID <- "01_206_143839A"
ClusterNumbToRmv="group2"
ClusterReorder <- c("group4","group5","group1","group3")

## 4th patients, 01_213_143839A
HierarchialClusterInfoFile <- paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_213_143839A_7clu.rds")
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_RmvCellByGiniMeanLowHighState_Kmean_01_213_143839A.rds")
MySampleID <- "01_213_143839A"
ClusterNumbToRmv=c("group2","group3","group4","group7")
ClusterReorder <- c("group1","group8","group6", "group5")

## 5th patients, 01_235_143921A
HierarchialClusterInfoFile <- paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_235_143921A_6clu.rds")
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_RmvCellByGiniMeanLowHighState_Kmean_01_235_143921A.rds")
MySampleID <- "01_235_143921A"
ClusterNumbToRmv=c("group3")
ClusterReorder <- c("group5","group1","group2", "group4","group6")

# ## 6th patients, 01_274_143921A
# HierarchialClusterInfoFile <- paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_274_143921A_6clu.rds")
# CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_RmvCellByGiniMean_01_274_143921A_LowHighState0.2.rds")
# MySampleID <- "01_274_143921A"
# ClusterNumbToRmv=c(1,6,5)
# ClusterReorder <- c(4,3,2)
# 
# ## 7th patients, 01_282_143855A
# HierarchialClusterInfoFile <- paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_282_143855A_7clu.rds")
# CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_RmvCellByGiniMean_01_282_143855A_LowHighState0.2.rds")
# MySampleID <- "01_282_143855A"
# ClusterNumbToRmv=c(1,6) # on Feb.11th, we decidied to remove cluster #6 
# ClusterReorder <- c(5,4,7,2,3)

## 8th patients, 09_025_143855A
HierarchialClusterInfoFile <- paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_09_025_143855A_5clu.rds")
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_RmvCellByGiniMeanLowHighState_Kmean_09_025_143855A.rds")
MySampleID <- "09_025_143855A"
ClusterNumbToRmv=c("group1","group4")
ClusterReorder <- c("group3","group5","group2")

## 9th patients, BM_CD138_143936A
HierarchialClusterInfoFile <- paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_BM_CD138_143936A_8clu.rds") ## group7 and group9 have cells < 10
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_RmvCellByGiniMeanLowHighState_Kmean_BM_CD138_143936A.rds")
MySampleID <- "BM_CD138_143936A"
ClusterNumbToRmv=c("group3","group4","group5") # group7 and group9 are already automatically removed.
ClusterReorder <- c("group2","group8","group6","group10","group1")

RemoveNormalClusterReorderHeatmap(HierarchialClusterInfoFile, CNVStateData_AfterRmvHighGiniMeanFile,MySampleID,ClusterNumbToRmv, ClusterReorder)

## It will generate
# "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_206_143839A.rds"
# CompHmapCNV_RmvByGiniMeanLowHighState_LowSt_AllChr_01_206_143839A_651cells_RordClone.pdf
# CompHmapCNV_RmvByGiniMeanLowHighState_LowSt_Chr1only_01_206_143839A_651cells_RordClone.pdf

## ClusterNumber_10LowCell_01_206_143839A_.txt.  Check this file and find the number of cluster which was remove. 

```

### Step4. Extract clone number for each cell. This is to split the merged .bam and run 
```{r Extract clone number for each cell} 
## 3rd patients, 01_206_143839A
HierarchialClusterInfoFile <- paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_206_143839A_6clu.rds")
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_RmvCellByGiniMean_01_206_143839A_LowHighState1.rds")
MySampleID <- "01_206_143839A"
ClusterNumbToRmv=6
ClusterReorder <- c(1,3,5,4,2)

## 8th patients, 09_025_143855A
HierarchialClusterInfoFile <- paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_09_025_143855A_5clu.rds")
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_RmvCellByGiniMean_01_206_143839A_LowHighState1.rds")
MySampleID <- "09_025_143855A"
ClusterNumbToRmv=c("group1","group4")
ClusterReorder <- c("group3","group5","group2")


RemoveNormalClusterReorder_CloneNumb <- function (HierarchialClusterInfoFile="RowSplitHeatmapDraw_WithoutArtifactLowHighState6_AllChr_01_213_143839A.rds",
                                      CNVStateData_AfterRmvHighGiniMeanFile="HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegion_01_213_143939A.rds",
                                       MySampleID,ClusterNumbToRmv=c(1,2), ClusterReorder=c(3,6,4,5)) {
      # HMMcopy_StateIdeal_NoRmvArtifactCell_ColNameProc <- readRDS(HMMcopyFile_BeforeRmvArtifact) ;
      # print(paste0("CNV state after removing by high Gini and Mean state: ", dim(HMMcopy_StateIdeal_NoRmvArtifactCell_ColNameProc))) # 881  5259

      if(is.null(CNVStateData_AfterRmvHighGiniMeanFile)) {
        HMMcopy_StateIdeal_AfterRmvArtifactCell_ColNameProc <- NULL
      } else {
        HMMcopy_StateIdeal_AfterRmvArtifactCell_ColNameProc <- readRDS(CNVStateData_AfterRmvHighGiniMeanFile)
      }
      dim(HMMcopy_StateIdeal_AfterRmvArtifactCell_ColNameProc) #  704cell 4922bins

      ########## ++++++++++ Remove artifacts by cluster number of cells, and reorder cells by cluster number
      HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster <-  RemoveReorderCell_ByClusterNumb(HMMcopy_StateIdeal_AfterRmvArtifactCell_ColNameProc,
                                                                   HierarchialClusterInfoFile, ClusterNumbToRmv, ClusterReorder, ClusterNumbToPick=NULL, MySampleID, ForPloidyPlot=FALSE)
      dim(HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster) #  651 cells  4922
      # saveRDS(HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster, file=paste0("HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_",MySampleID,".rds"))

      ##### Count the number of bins in each chromosome
      Position_chrStart <- grep("chr", colnames(HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster)); length(Position_chrStart) ; print(Position_chrStart) ## 22
      # [1]    1  416  877 1262 1629 1969 2298 2597 2834 3050 3300 3555 3808 3992 4163 4315 4457 4603 4747 4854 4969 5027 5089 5364
      EndChr1Position <- (Position_chrStart[2]-1); print(EndChr1Position) # 326
      EndChr1PositionPercent <- (Position_chrStart[2]-1)/ncol(HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster); print(EndChr1PositionPercent)

      # EndChrYPosition <- Position_chrStart[24]-1; print(EndChrYPosition) # 5363
      # EndChrYPositionPercent <- (Position_chrStart[24]-1)/ncol(HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster); print(EndChrYPositionPercent)

      ##### Count the number of cells in each cluster.
      NumberCell_EachCluster <- CountCell_EachCluster(HierarchialClusterInfoFile, ClusterReorder) # [1] 118  31 135  92  11
      Numb_Cell <- sum(NumberCell_EachCluster)

      ## Split cells in hierarchical clustering
      if(length(ClusterReorder)>=10) {
            MySplit<-c(paste('clone0', c( rep(1:9, times=NumberCell_EachCluster[1:9])), sep="" ),
                       paste('clone',c(rep(10:length(ClusterReorder), times=NumberCell_EachCluster[10:length(ClusterReorder)] )), sep="" ) )
      } else {
            MySplit<-paste('clone', c( rep(1:length(ClusterReorder), times=NumberCell_EachCluster)) )
      }

      print(MySplit[1:10]) # [1] "clone 1" "clone 1" "clone 1" "clone 1" "clone 1" "clone 1" "clone 1" "clone 1" "clone 1" "clone 1"
      table(MySplit); sum(table(MySplit)) # 651 
      # clone 1 clone 2 clone 3 clone 4 clone 5 
      # 19     133      59     415      25 

      ### Map between CellID and CloneID
      CellID_CloneID_DF <- data.frame(CellID=rownames(HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster), CloneNumb=MySplit); dim(CellID_CloneID_DF)
      CellID_CloneID_DF <- CellID_CloneID_DF %>% dplyr::mutate(CellID=gsub("^X","", CellID)) %>% dplyr::mutate(CellID=gsub("_","-", CellID)) %>% dplyr::mutate(CloneNumb=gsub(" ","", CloneNumb))
      
      fwrite(CellID_CloneID_DF, file=paste0(VignetteDir, "CellID_CloneID_", MySampleID, "_ForSplitBam.txt"), col.names=TRUE, row.names=FALSE, sep="\t",quote=FALSE)
}


## Pt8. 09_205
HierarchialClusterInfoFile <- paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_09_025_143855A_5clu.rds")
CNVStateData_AfterRmvHighGiniMeanFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_RmvCellByGiniMean_09_025_143855A_LowHighState0.2.rds")
MySampleID <- "09_025_143855A"
ClusterNumbToRmv=c("group1","group4")
ClusterReorder <- c("group3","group5","group2")

RemoveNormalClusterReorder_CloneNumb(HierarchialClusterInfoFile, CNVStateData_AfterRmvHighGiniMeanFile, MySampleID,ClusterNumbToRmv, ClusterReorder) 




```


### Step5. Heatmap for chromosome 1 pericentromeric region
```{r chr1 centromere region heatmap} 
# ## Pt. #1
# HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File<-paste0(VignetteDir, "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_040_143929A.rds")
# HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_040_143929A.rds")
# MySampleID <- "01_040_143929A"
# Chr1Centromere_Heatmap(HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File, HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile, MySampleID)
# 
# ## Pt. #2
# HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File<-paste0(VignetteDir, "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_056_143929A.rds")
# HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_056_143929A.rds")
# MySampleID <- "01_056_143929A"
# Chr1Centromere_Heatmap(HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File, HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile, MySampleID)

## Pt. #3
HierarchialClusterInfoFile <- paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_206_143839A_5clu.rds")
HMMcopy_AfterRmvCell_WithoutNABinRegionFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_206_143839A.rds")
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGiniStateLowHighFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_RmvCellByGiniMean_01_206_143839A_LowHighState1.rds") # This file is from Step1 output.It has Bin Region
MySampleID <- "01_206_143839A" 

## Pt. #4
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File<-paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_213_143839A_7clu.rds")
HMMcopy_AfterRmvCell_WithoutNABinRegionFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_213_143839A.rds")
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGiniStateLowHighFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_RmvCellByGiniMean_01_213_143839A_LowHighState0.2.rds") # This file is from Step1 output.It has Bin Region
MySampleID <- "01_213_143839A"
Chr1Centromere_Heatmap(HierarchialClusterInfoFile, HMMcopy_AfterRmvCell_WithoutNABinRegionFile, HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGiniStateLowHighFile, MySampleID)

## Pt. #5
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File<-paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_235_143921A_6clu.rds")
HMMcopy_AfterRmvCell_WithoutNABinRegionFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_235_143921A.rds")
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGiniStateLowHighFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_RmvCellByGiniMean_01_235_143921A_LowHighState0.18.rds") # This file is from Step1 output.It has Bin Region
MySampleID <- "01_235_143921A"
Chr1Centromere_Heatmap(HierarchialClusterInfoFile, HMMcopy_AfterRmvCell_WithoutNABinRegionFile, HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGiniStateLowHighFile, MySampleID)

# ## Pt. #6
# HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File<-paste0(VignetteDir, "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_274_143921A.rds")
# HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_274_143921A.rds")
# MySampleID <- "01_274_143921A"
# Chr1Centromere_Heatmap(HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File, HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile, MySampleID)
# 
# ## Pt. #7 patient, 01_282_143855A
# HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File<-paste0(VignetteDir, "HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_01_282_143855A.rds")
# HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_282_143855A.rds")
# MySampleID <- "01_282_143855A"
# Chr1Centromere_Heatmap(HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File, HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_StateLowHighReorderFile, MySampleID)

## 8th patient, 09_025_143855A
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File<-paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_09_025_143855A_5clu.rds")
HMMcopy_AfterRmvCell_WithoutNABinRegionFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_09_025_143855A.rds")
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGiniStateLowHighFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_RmvCellByGiniMean_09_025_143855A_LowHighState0.2.rds") 
MySampleID <- "09_025_143855A" 
Chr1Centromere_Heatmap(HierarchialClusterInfoFile, HMMcopy_AfterRmvCell_WithoutNABinRegionFile, HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGiniStateLowHighFile, MySampleID)


## 9th patient, BM_CD138_143936A
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGini_File<-paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_BM_CD138_143936A_8clu.rds")
HMMcopy_AfterRmvCell_WithoutNABinRegionFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_BM_CD138_143936A.rds")
HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGiniStateLowHighFile <- paste0(VignetteDir, "HMMcopy_StateIdeal_RmvCellByGiniMean_BM_CD138_143936A_LowHighState0.25.rds")  
MySampleID <- "BM_CD138_143936A"
Chr1Centromere_Heatmap(HierarchialClusterInfoFile, HMMcopy_AfterRmvCell_WithoutNABinRegionFile, HMMcopy_StateIdeal_RmvArtifactCNV_ByMeanGiniStateLowHighFile, MySampleID)

### Output file name: CompHeatmap_Chr1CentromereRegion_01_235_143921A_OOOcells.pdf

```

### Step6. Ploidy plot for all chromosomes. 
```{r chr1 Ploidy plot} 
library(regioneR)
library(karyoploteR)
library(CopyNumberPlots)


DLPpHMMCopyOutDir <- "/Users/lees130/Library/CloudStorage/OneDrive-NYULangoneHealth/N07_DLPplus_scWGS/04a_DLPplus_1qGain_8ptData"
VignetteDir <- "/Users/lees130/Library/CloudStorage/OneDrive-NYULangoneHealth/N07_DLPplus_scWGS/09d_HMMCopy_PCAKMeanClustering_Rpackage/HMMCopyPCAKMeanPloidy/vignettes"

# ### Patient #1 - 01_040_143929A
# CNVSegmentFile=paste0(DLPpHMMCopyOutDir,"/01_040_143929A/hmmcopy_segments.csv.gz");
# HierarchialClusterInfoFile=paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifact_AllChr_01_040_143929A.rds");
# HierarchialClusterInfoAfRmvClusterFile=paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_040_143929A_5clu.rds");
# # HierarchialClusterInfoAfRmvClusterChr1File=paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifact_OnlyChr1_ByClusterSize6_01_040_143929A.rds");
# CNVStateData_AfterRmvHighGiniMeanFile=paste0(VignetteDir, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_040_143929A.rds" );   #"/HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_040_143929A.rds");
# ClusterNumbToRmv=NULL;  ClusterNumbToPick=NULL; MySampleID="01_040_143929A"
# 
# ### Patient #2 - 01_056_143929A
# CNVSegmentFile=paste0(DLPpHMMCopyOutDir,"/01_056_143929A/hmmcopy_segments.csv.gz");
# HierarchialClusterInfoFile=paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifact_AllChr_01_056_143929A.rds");
# HierarchialClusterInfoAfRmvClusterFile=paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_056_143929A_6clu.rds");
# # HierarchialClusterInfoAfRmvClusterChr1File=paste0(VignetteDir, "RowSplitHeatmapDraw_WithoutArtifact_OnlyChr1_ByClusterSize6_01_056_143929A.rds");
# CNVStateData_AfterRmvHighGiniMeanFile=paste0(VignetteDir, "HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_056_143929A.rds" );   #"/HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_056_143929A.rds");
# ClusterNumbToRmv=NULL;  ClusterNumbToPick=NULL; MySampleID="01_056_143929A"

### Patient #3 - 01_206_143839A
CNVSegmentFile=paste0(DLPpHMMCopyOutDir,"/01_206_143839A/hmmcopy_segments.csv.gz");
HierarchialClusterInfoAfRmvClusterFile=paste0(VignetteDir, "/RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_206_143839A_5clu.rds");
CNVStateData_AfterRmvHighGiniMeanFile=paste0(VignetteDir, "/HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_206_143839A.rds" );   #"/HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_206_143839A.rds");
ClusterNumbToRmv=NULL;  ClusterNumbToPick=NULL; MySampleID="01_206_143839A"

### Patient #4 - 01_213_143839A
CNVSegmentFile=paste0(DLPpHMMCopyOutDir,"/01_213_143839A/hmmcopy_segments.csv.gz");
HierarchialClusterInfoAfRmvClusterFile=paste0(VignetteDir, "/RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_213_143839A_7clu.rds");
CNVStateData_AfterRmvHighGiniMeanFile=paste0(VignetteDir, "/HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_213_143839A.rds" );   #"/HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_206_143839A.rds");
ClusterNumbToRmv=NULL;  ClusterNumbToPick=NULL; MySampleID="01_213_143839A"

### Patient #5 - 01_235_143921A
CNVSegmentFile=paste0(DLPpHMMCopyOutDir,"/01_235_143921A/hmmcopy_segments.csv.gz");
HierarchialClusterInfoAfRmvClusterFile=paste0(VignetteDir, "/RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_235_143921A_6clu.rds");
CNVStateData_AfterRmvHighGiniMeanFile=paste0(VignetteDir, "/HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_235_143921A.rds" );   #"/HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_206_143839A.rds");
ClusterNumbToRmv=NULL;  ClusterNumbToPick=NULL; MySampleID="01_235_143921A"

# ### Patient #6 - 01_274_143921A
# CNVSegmentFile=paste0(DLPpHMMCopyOutDir,"/01_274_143921A/hmmcopy_segments.csv.gz");
# HierarchialClusterInfoFile=paste0(VignetteDir, "/RowSplitHeatmapDraw_WithoutArtifact_AllChr_01_274_143921A.rds");
# HierarchialClusterInfoAfRmvClusterFile=paste0(VignetteDir, "/RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_274_143921A_6clu.rds");
# # HierarchialClusterInfoAfRmvClusterChr1File=paste0(VignetteDir, "/RowSplitHeatmapDraw_WithoutArtifact_OnlyChr1_ByClusterSize6_01_274_143921A.rds");
# CNVStateData_AfterRmvHighGiniMeanFile=paste0(VignetteDir, "/HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_274_143921A.rds" );   #"/HMMcopy_StateIdeal_RmvCellByGiniMean_WithoutBinRegionAnnot_01_274_143921A.rds");
# ClusterNumbToRmv=NULL;  ClusterNumbToPick=NULL; MySampleID="01_274_143921A"
# 
# ### Patient #7 - 01_282_143855A
# CNVSegmentFile=paste0(DLPpHMMCopyOutDir,"/01_282_143855A/hmmcopy_segments.csv.gz");
# HierarchialClusterInfoFile=paste0(VignetteDir, "/RowSplitHeatmapDraw_WithoutArtifact_AllChr_01_282_143855A.rds");
# HierarchialClusterInfoAfRmvClusterFile=paste0(VignetteDir, "/RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_01_282_143855A_7clu.rds");
# # HierarchialClusterInfoAfRmvClusterChr1File=paste0(VignetteDir, "/RowSplitHeatmapDraw_WithoutArtifact_OnlyChr1_ByClusterSize6_01_274_143921A.rds");
# CNVStateData_AfterRmvHighGiniMeanFile=paste0(VignetteDir, "/HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_01_282_143855A.rds" );   
# ClusterNumbToRmv=NULL;  ClusterNumbToPick=NULL; MySampleID="01_282_143855A"

### Patient #8 - 09_025_143855A
CNVSegmentFile=paste0(DLPpHMMCopyOutDir,"/09_025_143855A/hmmcopy_segments.csv.gz");
# HierarchialClusterInfoFile=paste0(VignetteDir, "/RowSplitHeatmapDraw_WithoutArtifact_AllChr_09_025_143855A.rds");
HierarchialClusterInfoAfRmvClusterFile=paste0(VignetteDir, "/RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_09_025_143855A_5clu.rds");
# HierarchialClusterInfoAfRmvClusterChr1File=paste0(VignetteDir, "/RowSplitHeatmapDraw_WithoutArtifact_OnlyChr1_ByClusterSize6_01_274_143921A.rds");
CNVStateData_AfterRmvHighGiniMeanFile=paste0(VignetteDir, "/HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_09_025_143855A.rds" );   
ClusterNumbToRmv=NULL;  ClusterNumbToPick=NULL; MySampleID="09_025_143855A"


### Patient #9 - BM_CD138_143936A
CNVSegmentFile=paste0(DLPpHMMCopyOutDir,"/BM_CD138_143936A/hmmcopy_segments.csv.gz");
HierarchialClusterInfoAfRmvClusterFile=paste0(VignetteDir, "/RowSplitHeatmapDraw_WithoutArtifactLowHighState_AllChr_BM_CD138_143936A_8clu.rds");
CNVStateData_AfterRmvHighGiniMeanFile=paste0(VignetteDir, "/HMMcopy_StateIdeal_AfterRmvArtifactCell_RmvReorderCluster_BM_CD138_143936A.rds" );   
ClusterNumbToRmv=NULL;  ClusterNumbToPick=NULL; MySampleID="BM_CD138_143936A"



library(regioneR)
library(karyoploteR)
library(CopyNumberPlots)
data.panel <-1

Karyoplote_BySegment(CNVSegmentFile=CNVSegmentFile,# HierarchialClusterInfoFile=HierarchialClusterInfoFile,
                     HierarchialClusterInfoAfRmvClusterFile=HierarchialClusterInfoAfRmvClusterFile,HierarchialClusterInfoAfRmvClusterChr1File=NULL,
                     CNVStateData_AfterRmvHighGiniMeanFile,
                     CNVStateData_AfterRmvHighGiniMeanClusterFile=NULL,
                     ClusterNumbToRmv,  ClusterNumbToPick, ClusterNumbToPick_InChr1=ClusterNumbToPick_InChr1,MySampleID=MySampleID )






```


### Step7. Subclone phylogenetic tree
```{r phylogenetic tree}
## Pt3, 01_206  clone evolution
TreeData <- data.frame("sample" = c("01_206","01_206","01_206","01_206","01_206"), #name of sample, useful to read in all samples at once if processing multiple
                           "from" = c("x",   "Clone1", "x",      "Clone3","Clone4"), #parental Clones. 'x' is a virtual normal. 
                           "to"= c("Clone1", "Clone2", "Clone3", "Clone4","Clone5"))  #child Clone
                           #"score" = c() #a metric of CNV burden 
# Pt3 01_206 cell numbers 
NumbCellData <- data.frame("sample" = c("01_206","01_206", "01_206", "01_206","01_206","01_206"), #name of sample, useful to read in all samples at once if processing multiple
                                  "name" = c("x", "Clone1", "Clone2", "Clone3", "Clone4", "Clone5"), #Clone
                                  "counts"= c(0,19, 133,59,415, 25)) #number of cells in Clone
MySampleID <- "01_206";


## Make Phylogenetic tree
PhylogeneticTree(df=TreeData, num_cells_in=NumbCellData, MySampleID=MySampleID) 

## This will output "01_206_PhylogeneticTree.pdf" file. 

```
